Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene

### 2018-10-25 paper/paper.tex
Apply changes by Park

% \makeatletter
% \renewcommand*{\fnum@figure}{{\normalfont\bfseries \figurename~\thefigure}}
% \renewcommand*{\@caption@fignum@sep}{\textbf{ \textbar }}
% \makeatother

@@

%%%%%%%%%%%%%%%%%%%%
%Main
%%%%%%%%%%%%%%%%%%%%
\begin{document}
\clearpage{}
\begin{abstract}
  {\normalsize
  Atomically thin graphene nanopores is emerging as one of the most
  promising candidates for next-generation membrane technology owing
  to the ultrahigh permeability. However, the size-exclusive mechanism
  of selectivity limits its scalability as it remains technically
  challenging to drill subnanometer pores over a large area. Here we
  report macroscopic salt rejection through atomically-thin porous
  graphene with the pore sizes of 20$\pm$10 nanometers, overcoming the
  pore size limitation set by the hydrated radii. By electrostatically
  gating a sheet of porous graphene, we report a considerable
  reduction of salt flux by up to 55\% and 85\% with the gate voltage,
  based on our experiments and theoretical predictions,
  respectively. We systematically investigate the effects of salt
  concentration and species, including developing a theory to model
  the electrolyte diffusion through a nanopore in a sheet of gated
  graphene. The interplay between graphene quantum capacitance and
  electrical double layer is identified to be the main mechanism
  responsible for the observed salt rejection, when the pore size is
  comparable to the Debye screening length. Our findings reveal a new
  degree of freedom controlling the membrane potential and electrolyte
  permeation through porous two-dimensional materials.
  }
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
A central objective in the emerging field of nanofluidics is to
manipulate transport phenomena at nanometer scale to enable new
technological opportunities. The reduction of characteristic
dimensions in nanofluidics gives rise to substantially different
transport properties compared to their bulk counterparts
\cite{Schoch_2008}. In nanopores, for example, the effect of
surface-mediated transport becomes dominant when the pore size is
comparable to the solute diffusion length, enabling new applications,
including ionic diodes \cite{Karnik_2007}, field-effect transistors
\cite{Nam_2009}, desalination \cite{Heiranian_2015}, and
nanopore-based DNA sequencing \cite{Heerema_2016,Garaj_2013}, Porous,
atomically thin two-dimensional (2D) materials, such as graphene, are
emerging as ideal membrane materials due to their ultimate
permeability
\cite{Suk_2010,Jiang_2009,Celebi_2014,Koenig_2012,Drahushuk_2012}. Early
research has suggested that using nanoporous graphene as the
desalination membrane enables excellent salt rejection by preventing
the permeation of hydrated ions with a high water permeability
simultaneous
\cite{Cohen_Tanugi_2012,Suk_2014,Cohen_Tanugi_2014,Cohen_Tanugi_2015,O_Hern_2015,Surwade_2015}. Nevertheless,
the fabrication of subnanometer porous 2D membranes over a large area
with atomic precision remains technically challenging
\cite{Suk_2014,Rollings_2016,O_Hern_2012,Wang_2017}. Moreover, a small
portion of large pores can significantly diminish salt rejection.
Very recently, a number of reports have explored the effect of surface
charges on the selectivity of ion transport through graphene
nanopores, as a consequence of the electrostatic interactions between
solvated ions and the charged functional groups attached on graphene
\cite{Rollings_2016,Surwade_2014}. However, it remains controversial
that if the surface charges can result in a degree of salt rejection,
as the membrane charge and potential cannot be precisely quantified
and are sensitive to the surroundings and chemical treatment
\cite{Li_2008}. The difficulty in turn hinders fundamental
understanding of the interactions between the atomically thin 2D
membrane and electrolyte solution. Here, we demonstrate macroscopic
salt rejection through a sheet of large-area porous graphene using
electrostatic gating, thereby deepening the understanding of
ion-graphene interaction and its implications on nanoporous
graphene-based desalination applications.

@@

\Fig{} \ref{fig:1}a depicts the experimental setup characterizing
ionic diffusion through a sheet of double layer porous graphene
(PG). Two reservoirs, namely the high-concentration reservoir (HCR)
containing electrolyte solution with molar concentration $c_0$, and
the low-concentration reservoir (LCR) containing de-ionized (DI)
water, are separated by a PG membrane supported by polycarbonate
track-etched (PCTE) film. Magnetic stirrers are used to minimize the
effects of external concentration polarization, and a conductivity
probe is placed in the LCR to monitor the ionic concentration as a
function of time. The membrane is attached to a piece of copper tape
connected to a voltage source applying an electrical bias
$V_{\mathrm{G}}$, with the LCR grounded (\Fig{} \ref{fig:1}a
inset). We fabricated the PCTE-supported PG using the protocol
developed by our group \cite{Choi_2018}, as schematically shown in
\Fig{} \ref{fig:1}b. The normally distributed, circular pores with
pore diameters of 20$\pm$10 nm (\Fig{} \ref{fig:1}c inset) over a
large area were drilled on graphene.  By optimizing the subsequent
transfer process, a high surface coverage ($>$ 98\%) of graphene on
PCTE was reached (\Fig{} \ref{fig:1}c), allowing to reliably
characterize the ion transport behavior (details see Supplementary
Information Section S1).  The system presented here allows us to
systematically investigate the diffusive flux across a sheet of porous
graphene, $\boldsymbol{J}$, driven by the gradient of electrochemical
potential in the electrolyte medium.

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO4$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity versus time profile
(e.g. \Fig{} \ref{fig:2}a) remains linear within the measurement
duration (1 h), suggesting the concentration gradient across the
membrane remains constant. The diffusive fluxes
correspond to different salts were obtained by converting the
conductivity-time profiles with the calibration curves (blue bars in
\Fig{} \ref{fig:2}b and Supplementary \Fig{} S2). For comparison
purposes, based on the assumption that the membrane consists
cylindrical pores with uniform diameter and a low tortuosity $\tau$ of
$\sim{}$1.2 \cite{O_Hern_2012}, the molar diffusive flux through the
bare PCTE membrane, $\boldsymbol{J}_{\mathrm{PCTE}}$, follows:
  \begin{equation}
    \label{eq:j-pcte}
    \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi a^{2} N_{\mathrm{p}} D \Delta c}{\delta \tau}
  \end{equation}
where $a$ is the PCTE pore radius, $N_{\mathrm{p}}$ is numbers of pore
per area, $D$ is the salt diffusivity, and $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$, $\delta$ is the thickness of the PCTE membrane,
and $\tau$ is the tortuosity. By using the salt diffusivity values
provided by the PTCE vendor (Supplementary Table S1), the
calculated diffusive fluxes (red bars in \Fig{} \ref{fig:2}b) show
reasonable agreement with the experimental values.

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
to those of bare PCTE membrane, suggesting that the mass transfer
resistance of PG is not significant, as the permeability is high. We
observed a more pronounced decrease of diffusive flux in KCl,
K$_{2}$SO$_{4}$ and K$_{3}$[Fe(CN)$_{6}$] systems but not for the rest
of the salts. We hypothesized that the strong cation-$\pi$ interactions
between K$^{+}$ cations and graphene slow down the migration of cations
along the surface-mediated transport pathway \cite{Sun_2014}.

@@

We next discuss the effects of electrostatic gating on graphene. In
order to avoid any interference from the conductivity probe to the
membrane potential, the “three-interval” method, which had been used
in controlling the membrane potential in mesoporous carbon membranes
\cite{Surwade_2014}, is adopted here. \Fig{} \ref{fig:2}a presents the
representative data of the measured conductivity with respect to
time. Specifically, in the first interval, the gate voltage source is
turned off and the conductivity probe is on, allowing us to obtain the
conductivity-time profile corresponding to the diffusive flux through
the PG-PCTE membrane, giving the average slope $s_{1}$. In the second
interval, a voltage $V_{\mathrm{G}}$ is applied to the membrane after
turning off the conductivity probe, followed by the third interval, in
which the conductivity probe is turned on again to give the
corresponding slope, $s_{3}$. The slope corresponding to the second
interval, $s_{2}$, as the conductivity probe is off, is determined by
linearly interpolating the conductivities from the endpoint of
interval 1 to the starting point of interval 3. Accordingly, we define
the salt rejection factor, $\xi$, which is effectively equivalent to
the relative degree of diffusive flux decrease as follows:
\begin{equation}
  \label{eq:rejection}
  \xi = \frac{\bar{s} - s_{2}}{\bar{s}} = \frac{\boldsymbol{J}_{\mathrm{PG0}}
    - \boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})}{\boldsymbol{J}_{\mathrm{PG0}}}
\end{equation}
where $\bar{s}$ is the average slope of $s_{1}$ and $s_{3}$, namely
$(s_{1} + s_{3})/2$, and
$\boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})$ is the diffusive flux
across the PTCE-supported PG double layer as a function of
$V_{\mathrm{G}}$. Note that a positive $\xi$ between 0 and 1
represents a decrease of the diffusive flux. The factor $\xi$ enables
a reliable and stable measurement of salt rejection, as the salt
species and initial conditions may induce a degree of measurement
uncertainty between different samples. A control experiment was
performed to ensure the copper tape was not in direct contact with the
ionic solution (Supplementary Section 3).

Using 0.1 mM KCl solution in HCR, we measured $\xi$ as a function of
$V_{\mathrm{G}}$ within $\pm1.25$ V, before triggering the
electrochemical reactions (\Fig{} \ref{fig:2}c). We observe an
asymmetric dependence of $\xi$ with respect to $V_{\mathrm{G}}$, with
a higher degree of salt rejection for $V_{\mathrm{G}}>0$. where
positive carriers (holes) are induced in graphene. We notice that an
increase $\xi$, or a decrease of $\boldsymbol{J}_{\mathrm{PG}}$ with
$V_{\mathrm{G}}$, showing an inverse trend compared to those observed
in ionic field effect transistors (IFETs), in which the diffusive flux
increases with the gate voltage \cite{Nam_2009,Cheng_2018}. A further
increase of the KCl concentration $c_{0}$ influences the attainable
degree of salt rejection. $\xi$ values measured at
$V_{\mathrm{G}}$ = 1.25 V, $\xi_{\mathrm{max}}$, as a function of
$c_{0}$ from 10$^{-4}$ to 10$^{-2}$ M KCl (see Supplementary Section
S4) exhibit a power law dependency (\Fig{} \ref{fig:2}d). The
$\xi_{\mathrm{max}}-c_{0}$ relation can be nicely fitted by a power
law function following
$\xi_{\mathrm{max}} c_{0}^{0.59} = \mathrm{Constant}$. As the Debye
screening length in solution, $\lambda_{\mathrm{D}}$, is inversely
proportional to $c^{0.5}$, we infer that the $V_{\mathrm{G}}$-induced
salt rejection originates from the modulation of the electrical double
layer (EDL), as will be discussed later.

@@

$\psi_{\mathrm{G}}$ tp $\sim$0.3 V at $V_{\mathrm{G}}$ = 1.25 V,

@@

\subsection{Salt rejection mechanism}
\label{sec:mechanism}
The numerical procedure described above allows us to resolve the
concentration and electric potential profiles near to a graphene
nanopore for a given $V_{\mathrm{G}}$. Since the ionic flux is driven
by both $\psi$ and $c$ fields (see Eq. \eqref{eq:pnp} and \eqref{eq:mu}),
we focus on the electrochemical potential $\mu_{i}$, which represents
the combined driving force, to reveal its dependence on the applied
$V_{\mathrm{G}}$. Following the same system considered in \Fig{}
\ref{fig:3}b, the calculated axisymmetric electric potential $\psi$
and the relative electrochemical potentials,
$\Delta \mu_{\mathrm{K^{+}}}$ and $\Delta \mu_{\mathrm{Cl^{-}}}$, for
$V_{\mathrm{G}}$ = 0.75 V are shown in \Fig{} \ref{fig:4}a and
\ref{fig:4}b, respectively. The reference point of the electrochemical
potentials is set at the bulk phase in the HCR. As expected, the
electric potential reaches the maximum at the graphene surface (with
$\psi_{\mathrm{G}}$ of $\sim$100 mV) and decays towards the bulk
solution phase, forming an EDL surrounding the graphene surface. Since
the Debye screening length $\lambda_{\mathrm{D}}$ is comparable to the
pore radius $R_{\mathrm{G}}$, the electric potential at the nanopore
center remains at $\sim$25 mV, comparable to the thermal energy at
room temperature ($k_{\mathrm{B}}T$ = 26 meV). Consequently, it is
evident that the potential barrier is sufficient to modulate the
diffusive flux.

@@

Finally, we examine other salt species and compare the degree of salt
rejection between experiments and simulations. \Fig{} \ref{fig:6}a
presents the measured salt rejection $\xi$ as a function of
$V_{\mathrm{G}}$ at $c_{0}$ = 0.1 mM, for the seven salt species
considered here, including NaCl, LiCl, KCl, MgSO$_{4}$, CaCl$_{2}$,
K$_{2}$SO$_{4}$, and K$_{3}$[Fe(CN)$_{6}$] (Supplementary
Section S7). Similar to the KCl system discussed earlier, the salt
rejection is more pronounced in the positive $V_{\mathrm{G}}$
regime. The degree of salt rejection deceases with the salt valance;
for example, the measured $\xi$ values at $V_{\mathrm{G}}$ = 1.25 V
decreases from $>$ 0.5 for monovalent NaCl to -0.04 for multivalent
K$_{3}$[Fe(CN)$_{6}$]. This observation is endorsed by our simulations,
in which we calculate $\xi$ as a function of $V_{\mathrm{g}}$ using
the same simulation setups(\Fig{} \ref{fig:6}b). The calculated
$\xi_{\mathrm{max}}$ values are quantitatively compared with
experiments (\Fig{} \ref{fig:6}c), and the same trend is
observed. Indeed, because the Debye screening length
$\lambda_{\mathrm{D}}$ decreases with the salt valence, the graphene
surface potential appears to decrease in the multivalent salt systems,
thereby reducing $\xi$ (see \Fig{} \ref{fig:5}b). The calculated and
experimental $\xi_{\mathrm{max}}$ as a function of
$\lambda_{\mathrm{D}}$ are consequently plotted, for the seven salt
species considered here (\Fig{} \ref{fig:6}d). We notice that without
any fitting parameters, our theory can predict the experimental values
reasonably well. In addition, the
$\xi_{\mathrm{max}}-\lambda_{\mathrm{D}}$ dependence is approximately
linear, suggesting that the effect of salt diffusivity is
negligible. This finding also explains the concentration dependence of
$\xi_{\mathrm{max}}$ observed in \Fig{} \ref{fig:2}d, as
$\lambda_{\mathrm{D}} \propto c_{0}^{-0.5}$.

In summary, we have presented a comparative experimental and modeling
study on macroscopic salt rejection through a sheet of large-area
porous graphene under electrostatic gating. We show that due to a
small quantum capacitance of atomically-thin graphene, the graphene
surface potential is considerably lower than the applied gate
voltage. As a result, the subtle redistribution of the electrochemical
potentials create new pathways for ionic transport, which in turn lead
to a considerably degree of salt rejection. We report a high degree of
salt rejection at $V_{\mathrm{G}}$ = 1.25 V of up to 55\% and 85\%,
based on our experiments and theoretical predictions, respectively. We
demonstrate that the observed salt rejection positively correlates
with the Debye screening length that nonlinearly modulates the
graphene surface potential and charge density. The experimental
results and fundamental principles presented here open an avenue
towards realization of atomically thin 2D porous membranes for
advanced separation process in electrolyte solutions.

@@

\begin{acknowledgement}
\todo[inline]{TBD} R.M.W. and H.G.P. acknowledge the support from
XXX. T.T. and C.J.S. are grateful for the financial support from the
ETH startup funding.
\end{acknowledgement}

>>>>

@@

%%%%%%%%%%%%%%%%%%%%
%Main
%%%%%%%%%%%%%%%%%%%%
\begin{document}
\clearpage{}
\begin{abstract}
  {\normalsize Atomically thin graphene nanopores is emerging as one
    of the most promising candidates for next-generation membrane
    technology owing to the ultrahigh permeability. However, the
    size-exclusive mechanism of selectivity limits its scalability as
    it remains technically challenging to drill subnanometer pores
    over a large area. Here we report macroscopic salt rejection
    through atomically-thin porous graphene with the pore sizes of
    20$\pm$10 nanometers, and a pore density of 1.25$\times$10$^{10}$
    cm$^{-2}$, overcoming the pore size limitation set by the hydrated
    radii. By electrostatically gating a sheet of porous graphene, we
    report a considerable reduction of salt flux by up to 55\% and
    85\% with the gate voltage, based on our experiments and
    theoretical predictions, respectively. We systematically
    investigate the effects of salt concentration and species,
    including developing a theory to model the electrolyte diffusion
    through a nanopore in a sheet of gated graphene. The interplay
    between graphene quantum capacitance and electrical double layer
    is identified to be the main mechanism responsible for the
    observed salt rejection, when the pore size is comparable to the
    Debye length. Our findings reveal a new degree of
    freedom controlling the membrane potential and electrolyte
    permeation through porous two-dimensional materials.  }
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
A central objective in the emerging field of nanofluidics is to
manipulate transport phenomena at nanometer scale to enable new
technological opportunities. The reduction of characteristic
dimensions in nanofluidics gives rise to substantially different
transport properties in comparison to their bulk counterparts
\cite{Schoch_2008}. In nanopores, for example, the effect of
surface-mediated transport becomes dominant when the pore size is
comparable to the characteristic length of diffusion, enabling new
applications such as ionic diodes \cite{Karnik_2007}, field-effect
transistors \cite{Nam_2009}, desalination \cite{Heiranian_2015}, and
nanopore-based DNA sequencing \cite{Heerema_2016,Garaj_2013}, Porous,
atomically thin two-dimensional (2D) materials, such as graphene, are
emerging as ideal membrane materials due to their ultimate permeation
\cite{Suk_2010,Jiang_2009,Celebi_2014,Koenig_2012,Drahushuk_2012}. Early
research has suggested that using nanoporous graphene as the
desalination membrane enables excellent salt rejection by preventing
the permeation of hydrated ions with a high water permeability
simultaneous
\cite{Cohen_Tanugi_2012,Suk_2014,Cohen_Tanugi_2014,Cohen_Tanugi_2015,O_Hern_2014,O_Hern_2015,Surwade_2015,Walker_2017,Ghosh_2018}. Nevertheless,
the fabrication of subnanometer porous 2D membranes over a large area
with atomic precision remains technically challenging
\cite{Suk_2014,Rollings_2016,O_Hern_2012,Wang_2017}. Moreover, a small
portion of excessively large pores can significantly diminish salt
rejection.  Very recently, a number of reports have explored the
effect of surface charges on the selectivity of ion transport through
graphene nanopores, as a consequence of the electrostatic interactions
between solvated ions and the charged functional groups attached on
graphene \cite{Rollings_2016,Surwade_2014}. However, it remains
controversial that if the surface charges can result in a degree of
salt rejection, as the membrane charge and potential cannot be
precisely quantified and are sensitive to the surroundings and
chemical treatment \cite{Li_2008}. The difficulty in turn hinders a
fundamental understanding of the interactions between the atomically
thin 2D membrane and the electrolyte solution. Here, we demonstrate
macroscopic salt rejection through a sheet of large-area porous
graphene using electrostatic gating, thereby deepening the
understanding of ion-graphene interaction and its implications on
nanoporous graphene-based desalination applications.

@@

\Fig{} \ref{fig:1}a depicts the experimental setup characterizing
ionic diffusion through a sheet of double layer porous graphene
(PG). Two reservoirs, namely the high-concentration reservoir (HCR)
containing electrolyte solution with molar concentration $c_0$, and
the low-concentration reservoir (LCR) containing de-ionized (DI)
water, are separated by a PG membrane supported by polycarbonate
track-etched (PCTE) film. Magnetic stirrers are used to minimize the
effects of external concentration polarization, and a conductivity
probe is placed in the LCR to monitor the ionic concentration as a
function of time. The membrane is attached to a piece of copper tape
connected to a voltage source applying an electrical bias
$V_{\mathrm{G}}$, with the LCR grounded (\Fig{} \ref{fig:1}a
inset). We fabricated the PCTE-supported PG using the protocol
developed by our group \cite{Choi_2018}, as schematically shown in
\Fig{} \ref{fig:1}b. The normally distributed, circular pores with
pore diameters of 20$\pm$10 nm (\Fig{} \ref{fig:1}c inset) over a
large area were perforated on graphene.  By optimizing the subsequent
transfer process, a high surface coverage ($>$ 98\%) of PG on PCTE was
obtained (\Fig{} \ref{fig:1}c), allowing reliable characterization of
the ion transport behavior (details see Supplementary Information
Section S1).  The system presented here allows us to systematically
investigate the diffusive flux of ions across a sheet of PG,
$\boldsymbol{J}$, driven by the gradient of electrochemical potential
in the electrolyte medium.

First, the control experiments were carried out by measuring the ionic
diffusion of seven salt species, including KCl, NaCl, LiCl,
K$_{2}$SO$_{4}$, MgSO4$_{4}$, CaCl$_{2}$ and K$_{3}$[Fe(CN)$_{6}$],
through the bare PCTE membrane at $c_{0}$ = 0.1 mM. As the diffusive
flux is comparably small, the conductivity varies linearly with time
(e.g. \Fig{} \ref{fig:2}a) within the measurement duration (1 h),
suggesting the concentration gradient across the membrane remains
constant. The diffusive fluxes of different salts were
obtained by converting the conductivity-time profiles with the
calibration curves (blue bars in \Fig{} \ref{fig:2}b and Supplementary
\Fig{} S2). For comparison purposes, based on the assumption that the
membrane consists cylindrical pores with uniform diameter and a low
tortuosity $\tau$ of $\sim{}$1.2 \cite{O_Hern_2012}, the molar
diffusive flux through the bare PCTE membrane,
$\boldsymbol{J}_{\mathrm{PCTE}}$, follows:
  \begin{equation}
    \label{eq:j-pcte}
    \boldsymbol{J}_{\mathrm{PCTE}} = \frac{\pi a^{2} N_{\mathrm{p}} D \Delta c}{\delta \tau}
  \end{equation}
where $a$ is the PCTE pore radius, $N_{\mathrm{p}}$ is numbers of pore
per area, $D$ is the salt diffusivity, and $\Delta c$ is the bulk
concentration difference between two reservoirs, i.e.,
$\Delta c \sim c_{0}$, $\delta$ is the thickness of the PCTE membrane,
and $\tau$ is the tortuosity of the PCTE membrane. By using the salt diffusivity values
provided by the PTCE vendor (Supplementary Table S1), the
calculated diffusive fluxes (red bars in \Fig{} \ref{fig:2}b) show
reasonable agreement with the experimental values.

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
with those of bare PCTE membrane, suggesting that the mass transfer
resistance of PG is not significant, as the permeability is high. We
observed a more pronounced decrease of diffusive flux in KCl,
K$_{2}$SO$_{4}$ and K$_{3}$[Fe(CN)$_{6}$] systems but not for the rest
of the salts. We hypothesized that the strong cation-$\pi$ interactions
between K$^{+}$ cations and graphene slow down the migration of cations
along the surface-mediated transport pathway \cite{Sun_2014}.

@@

We next discuss the effects of electrostatic gating on graphene. In
order to avoid any interference from the conductivity probe to the
membrane potential, the “three-interval” method, which has been used
in controlling the membrane potential in mesoporous carbon membranes
\cite{Surwade_2014}, is adopted here. \Fig{} \ref{fig:2}a
illustratively represents how the measured conductivity evolves with
time. Specifically, in the first interval, the gate voltage source is
turned off and the conductivity probe is on, allowing us to obtain the
conductivity-time profile with an average slope $s_{1}$ that
corresponds to the diffusive flux across the PG-PCTE membrane, . In
the second interval, a voltage $V_{\mathrm{G}}$ is applied to the
membrane after turning off the conductivity probe, followed by the
third interval $s_{3}$, in which the conductivity probe is turned on
again to give the corresponding slope. The slope corresponding to the
second interval $s_{2}$, as the conductivity probe is off, is
determined by linear interpolation of the conductivities from the
endpoint of interval 1 to the starting point of interval
3. Accordingly, we define the salt rejection factor, $\xi$, which is
effectively equivalent to the relative degree of diffusive flux
decrease as follows:
\begin{equation}
  \label{eq:rejection}
  \xi = \frac{\bar{s} - s_{2}}{\bar{s}} = \frac{\boldsymbol{J}_{\mathrm{PG0}}
    - \boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})}{\boldsymbol{J}_{\mathrm{PG0}}}
\end{equation}
where $\bar{s}$ is the average slope of $s_{1}$ and $s_{3}$, namely
$(s_{1} + s_{3})/2$, and
$\boldsymbol{J}_{\mathrm{PG}}(V_{\mathrm{G}})$ is the diffusive flux
across the PTCE-supported PG double layer as a function of
$V_{\mathrm{G}}$. Note that a positive $\xi$ between 0 and 1
represents a decrease of the diffusive flux. The factor $\xi$ enables
a reliable and stable measurement of salt rejection, as the salt
species and initial conditions may induce a degree of measurement
uncertainty between different samples. A control experiment was
performed to ensure the copper tape was not in direct contact with the
ionic solution (Supplementary Section 3).

Using 0.1 mM KCl solution in HCR, we measured $\xi$ as a function of
$V_{\mathrm{G}}$ within $\pm1.25$ V, before triggering the
electrochemical reactions (\Fig{} \ref{fig:2}c). We observe an
asymmetric dependence of $\xi$ with respect to $V_{\mathrm{G}}$, with
a higher degree of salt rejection for $V_{\mathrm{G}}>0$. where
positive carriers (holes) are induced in graphene. We notice that an
increase of $\xi$, or a decrease of $\boldsymbol{J}_{\mathrm{PG}}$
with $V_{\mathrm{G}}$, showing an inverse trend compared with those
observed in ionic field effect transistors (IFETs) in which the
diffusive flux increases with the gate voltage
\cite{Nam_2009,Cheng_2018}. A further increase of the KCl
concentration $c_{0}$ influences the attainable degree of salt
rejection. $\xi$ values measured at $V_{\mathrm{G}}$ = 1.25 V,
$\xi_{\mathrm{max}}$, as a function of $c_{0}$ from 10$^{-4}$ to
10$^{-2}$ M KCl (see Supplementary Section S4) exhibit a power law
dependency (\Fig{} \ref{fig:2}d). The $\xi_{\mathrm{max}}-c_{0}$
relation can be nicely fitted by a power law function following
$\xi_{\mathrm{max}} c_{0}^{0.59} = \mathrm{constant}$. As the Debye
length in solution, $\lambda_{\mathrm{D}}$, is inversely proportional
to $c^{0.5}$, we infer that the $V_{\mathrm{G}}$-induced salt
rejection originates from the modulation of the electrical double
layer (EDL), as will be discussed later.

@@

$\psi_{\mathrm{G}}$ to $\sim$0.3 V at $V_{\mathrm{G}}$ = 1.25 V,

@@

\subsection{Salt rejection mechanism}
\label{sec:mechanism}
The numerical procedure described above allows us to resolve the
concentration and electric potential profiles near a graphene
nanopore for a given $V_{\mathrm{G}}$. Since the ionic flux is driven
by both $\psi$ and $c$ fields (see Eq. \eqref{eq:pnp} and \eqref{eq:mu}),
we focus on the electrochemical potential $\mu_{i}$, which represents
the combined driving force, to reveal its dependence on the applied
$V_{\mathrm{G}}$. Following the same system considered in \Fig{}
\ref{fig:3}b, the calculated axisymmetric electric potential $\psi$
and the relative electrochemical potentials,
$\Delta \mu_{\mathrm{K^{+}}}$ and $\Delta \mu_{\mathrm{Cl^{-}}}$, for
$V_{\mathrm{G}}$ = 0.75 V are shown in \Fig{} \ref{fig:4}a and
\ref{fig:4}b, respectively. The reference point of the electrochemical
potentials is set at the bulk phase in the HCR. As expected, the
electric potential reaches the maximum at the graphene surface (with
$\psi_{\mathrm{G}}$ of $\sim$100 mV) and decays towards the bulk
solution phase, forming an EDL surrounding the graphene surface. Since
the Debye length $\lambda_{\mathrm{D}}$ is comparable to the
pore radius $R_{\mathrm{G}}$, the electric potential at the nanopore
center remains at $\sim$25 mV, comparable to the thermal energy at
room temperature ($k_{\mathrm{B}}T$ = 26 meV). Consequently, it is
evident that the potential barrier is sufficient to modulate the
diffusive flux.

@@

Finally, we examine other salt species and compare the degree of salt
rejection between experiments and simulations. \Fig{} \ref{fig:6}a
presents the measured salt rejection $\xi$ as a function of
$V_{\mathrm{G}}$ at $c_{0}$ = 0.1 mM, for the seven salt species
considered here, including NaCl, LiCl, KCl, MgSO$_{4}$, CaCl$_{2}$,
K$_{2}$SO$_{4}$, and K$_{3}$[Fe(CN)$_{6}$] (Supplementary
Section S7). Similar to the KCl system discussed earlier, the salt
rejection is more pronounced in the positive $V_{\mathrm{G}}$
regime. The degree of salt rejection deceases with the salt valance;
for example, the measured $\xi$ values at $V_{\mathrm{G}}$ = 1.25 V
decreases from $>$ 0.5 for monovalent NaCl to -0.04 for multivalent
K$_{3}$[Fe(CN)$_{6}$]. This observation is endorsed by our simulations,
in which we calculate $\xi$ as a function of $V_{\mathrm{g}}$ using
the same simulation setups(\Fig{} \ref{fig:6}b). The calculated
$\xi_{\mathrm{max}}$ values are quantitatively compared with
experiments (\Fig{} \ref{fig:6}c), and the same trend is
observed. Indeed, because the Debye length
$\lambda_{\mathrm{D}}$ decreases with the salt valence, the graphene
surface potential appears to decrease in the multivalent salt systems,
thereby reducing $\xi$ (see \Fig{} \ref{fig:5}b). The calculated and
experimental $\xi_{\mathrm{max}}$ as a function of
$\lambda_{\mathrm{D}}$ are consequently plotted, for the seven salt
species considered here (\Fig{} \ref{fig:6}d). We notice that without
any fitting parameters, our theory can predict the experimental values
reasonably well. In addition, the
$\xi_{\mathrm{max}}-\lambda_{\mathrm{D}}$ dependence is approximately
linear, suggesting that the effect of salt diffusivity is
negligible. This finding also explains the concentration dependence of
$\xi_{\mathrm{max}}$ observed in \Fig{} \ref{fig:2}d, as
$\lambda_{\mathrm{D}} \propto c_{0}^{-0.5}$.

In summary, we have presented a comparative experimental and modeling
study on macroscopic salt rejection through a sheet of large-area
porous graphene under electrostatic gating. We show that due to a
small quantum capacitance of atomically-thin graphene, the graphene
surface potential is considerably lower than the applied gate
voltage. As a result, the subtle redistribution of the electrochemical
potentials create new pathways for ionic transport, which in turn lead
to a considerable degree of salt rejection. We report a high degree of
salt rejection at $V_{\mathrm{G}}$ = 1.25 V of up to 55\% and 85\%,
based on our experiments and theoretical predictions, respectively. We
demonstrate that the observed salt rejection positively correlates
with the Debye length that nonlinearly modulates the graphene surface
potential and charge density. The experimental results and fundamental
principles presented here open an avenue towards realization of
atomically thin 2D porous membranes for advanced separation process in
electrolyte solutions.

@@

\begin{acknowledgement}
  R.M.W. acknowledges support from the Binnig and Rohrer
  Nanotechnology Center of ETH Zurich and IBM Zurich. T.T. and
  C.J.S. are grateful for the financial support from the ETH startup
  funding.
\end{acknowledgement}
